---
title: "Analysis Belgian Start-Ups CHP"
author: "Joost Haddinga"
date: "5/6/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(dynlm)
library(AER)
library(lmtest)
library(sandwich)
library(zoo)
library(data.table)
```
Load data previously collected in dropbox
```{r}
IWT_database <- read_excel("~/Desktop/Bachelor_TiU/CHP/Copy of IWT Longitudinale database (without y1).xlsx")


IWT_data <- IWT_database %>% select(-ageinmonths) %>%
  rename(assets = 'Activa (assets, 20/58)') %>%
  rename(turnover = 'bedrijfsopbrengsten (turnover;revenues, 70/76A)') %>%
  rename(ebit = 'bedrijfswinst/verlies voor belastingen (profit/losses before tax, 9903)') %>%
  rename(gross_margin = 'brutomarge (gross margin, 9900)') %>%
  rename(net_income = 'winst/verlies van het boekjaar (net income, 9904)') %>%
  rename(employees = 'employees (code 100.3 or 1003) or 9087')%>%
  mutate(ebit = as.numeric(ebit))%>%
  mutate(net_income = as.numeric(net_income))
  
head(IWT_data)
```

Basic summary statistics
```{r}
summary(IWT_data)

summary(IWT_data %>% group_by(Survival_years))

summary(IWT_data %>% group_by(founding_year))


```

## Basic plots - levels
```{r}
ggplot(IWT_data, aes(Survival_years, assets, group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, turnover, group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, gross_margin, group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, net_income, group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, ebit, group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, employees, group=company_name))+geom_line()
```
## Basic plots - logs
```{r}
ggplot(IWT_data, aes(Survival_years, log(assets*1000), group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, log(turnover*1000), group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, log(gross_margin*1000), group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, log(net_income*1000), group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, log(ebit*1000), group=company_name))+geom_line()

ggplot(IWT_data, aes(Survival_years, employees, group=company_name))+geom_line()

#logs problematic whenever negative values or 0 - gross margin, ebit, net income
```
```{r}
ggplot(IWT_data, aes(Survival_years, log(assets*1000), group=company_name))+geom_line()+facet_wrap(~founding_year)

ggplot(IWT_data, aes(Survival_years, log(ebit*1000), group=company_name))+geom_line()+facet_wrap(~founding_year)

ggplot(IWT_data, aes(log(assets)))+geom_boxplot()+facet_wrap(~Survival_years)
```



## Data manipulation
  * Build controls for final status
  * Trim data, make it neat
  * Build controls for sector
  * Build controls for time
  * Build controls for negative values and developments
  * Get correlations
  
```{r}
IWT_data1 <- IWT_data %>%
  select(-turnover) %>%
  drop_na() %>%
  mutate(sector = ifelse(sector1==1, 1, ifelse(sector2==1, 2, ifelse(sector3==1, 3, ifelse(sector4 ==1, 4, ifelse(sector5==1,5,6)))))) %>%
  mutate(margin_growth = ifelse(Firstyear != 1, (gross_margin-shift(gross_margin)/shift(gross_margin)),0.001)) %>%
  mutate(income_growth = ifelse(Firstyear != 1, (net_income-shift(net_income)/shift(net_income)),0.001)) %>%
  mutate(ebit_growth = ifelse(Firstyear != 1, (ebit-shift(ebit))/shift(ebit),0.001))%>%
  mutate(lassets = log(assets)) %>% filter(lassets > -100000)

IWT_data2 <- IWT_data1 %>%
  group_by(company_name)%>%
  mutate(final_status = ifelse(failureTOT==1,'Failed', ifelse(acquired==1, 'Acquired', ifelse(merged==1, 'Merged', 'Self-sufficient'))))%>% #Make these transfer to all observations of status 
ungroup()



cor(IWT_data2$assets, IWT_data2$net_income, use="complete.obs")
cor(IWT_data2$lassets, IWT_data2$net_income, use="complete.obs") #Why not working?
cor(IWT_data2$lassets, IWT_data2$income_growth, use="complete.obs") #Why not working?
cor(IWT_data2$assets, IWT_data2$Survival_years, use="complete.obs")
cor(IWT_data2$net_income, IWT_data2$employees, use="complete.obs")
cor(IWT_data2$assets, IWT_data2$employees, use="complete.obs")
cor(IWT_data2$assets, IWT_data2$founding_year, use="complete.obs")
cor(IWT_data2$net_income, IWT_data2$margin_growth, use="complete.obs")
```

## Plots with manipulated data
  * Include various facets
  * Include relationships between variables (excl. time) - but maybe control for time
  * Format graphs for more readability
  * Include basic regressions
```{r}
ggplot(IWT_data2, aes(Survival_years, assets, group=company_name))+geom_line()+facet_wrap(~sector)


ggplot(IWT_data2, aes(Survival_years, log(assets), group=company_name))+geom_line()+facet_wrap(~sector)

ggplot(IWT_data2, aes(Survival_years, assets, group=company_name))+geom_line()+facet_wrap(~founding_year)


ggplot(IWT_data2, aes(Survival_years, log(assets), group=company_name))+geom_line()+facet_wrap(~founding_year)

ggplot(IWT_data2, aes(Survival_years, gross_margin, group=company_name))+geom_line()

ggplot(IWT_data2, aes(Survival_years, net_income, group=company_name))+geom_line()
#Problems with regressions: Different time paths due to different foundation years and different lengths of observation - additional to Survivorship bias

ggplot(IWT_data2, aes(Survival_years, ebit, group=company_name))+geom_line()

ggplot(IWT_data2, aes(Survival_years, employees, group=company_name))+geom_line()
```

```{r}
ggplot(IWT_data2, aes(Survival_years, income_growth, group=company_name))+geom_line()

ggplot(IWT_data2, aes(Survival_years, income_growth, group=company_name))+geom_line()+facet_wrap(~sector)

ggplot(IWT_data2, aes(Survival_years, ebit_growth, group=company_name))+geom_line() #Graph seems very odd

ggplot(IWT_data2, aes(Survival_years, margin_growth, group=company_name))+geom_line()

ggplot(IWT_data2, aes(Survival_years, margin_growth, group=company_name))+geom_line()+facet_wrap(~sector)
#smoothers not helpful due to different legths of trajectories
```


## Simple regressions
  * levels
  * logs
  * only specify assets
  * all specifications
  * test heterogeneity / endogeneity
  * include controls / lags
```{r}
#What to regress assets upon to symbolize growth of a company?
mod1 <- lm(net_income ~ assets + founding_year, data=IWT_data2)
summary(mod1)

mod2 <- lm(net_income ~ assets +employees + founding_year, data=IWT_data2)
summary(mod2)

mod3 <- lm(Survival_years ~ assets +employees + founding_year, data=IWT_data2)
summary(mod3)


mod4 <- lm(Survival_years ~ lassets +employees + founding_year, data=IWT_data2)
summary(mod4)

mod5 <- lm(Survival_years ~ lassets +employees + founding_year+income_growth+margin_growth, data=IWT_data2)
summary(mod5)
mod6 <- lm(Survival_years ~ lassets + founding_year+net_income+gross_margin, data=IWT_data2)
summary(mod6)
mod7 <- lm(Survival_years ~ lassets +employees + founding_year+sector1+sector2+sector3+sector4+sector5, data=IWT_data2)
summary(mod7)
```

